Real-space computation of $E$/$B$-mode maps I: Formalism and Compact Kernels

### 2018-06-20 articles/formalism/sections/kernel_paintbrush.tex
Updated kernel paintbrush discussion.

The $B$ part of the Green's function, traced along the filament, cancels itself except at the filament ends.  This results in a non-zero $B$ pattern for the filament.  For a North--South filament, the $B$-mode pattern is positive on the North-East and South-West, and negative in the North-West and South-East.

@@

The intuition from the real-space kernels holds also when we distort the shape of the filament.  If the filament were bent around into a circle, the positive and negative parts of the $B$ pattern will cancel, and we are left with a hoop of pure $E$ pattern. Here it is important to note that this cancellation will not happen for an ellipse or any loop of non-contant radius of curvature.  The same general description holds for a spiral-shaped filament, which can can be viewed as distortion of the straight filament.  The filament is highlighted by positive $E>0$.  The $E$-pattern is more negative on the interior of a curve than on the exterior, and the concentric rings of filamentary structure make an increasingly negative $E$ value inside. While the $B$-pattern is again concentrated at the ends of the filament in an oriented pair of positive/negative fluctuations, note that it is non-vanishing in the intermediate regions due to the varying radius of curvature of the spiral.
%\footnote{ \revisit{Why do we use this specific spiral - Its easy to compute the tangent to the spiral at any point.}}.

\comment{If this is correct, then one expects to see a higher B-mode signal in intermediate regions of the spiral, as compared to the B-mode signal in the intermediate regions of the linear filament. This  then should be more visible in a logarithmically scaled version of Fig. 8.}

\revisit{Note that we have made the simplifying assumption that the polarization headless vectors are perpendicular to the magnetic field which is always assumed parallel to the tangent to the filament at any given point. This results in generating dominantly an E-mode signal while the B-mode signal is present only near the edges. By similar arguments it can be shown that that polarization pattern will be dominated by B-modes if the polarization vectors were aligned at $+45$ (or $-45$) degrees to the tangent to the filament while the E-mode signal would only be present at the edges. One way to get this configuration is to change the orientation of the magnetic field vector with respect to the tangent to the filament. Therefore a careful study of the $E/B$ mode power in filaments can provide insights into the orientation of the magnetic field with respect to the length of the filament.}

%\revisit{A stacking analysis of Planck data \citep{2016A&A...586A.141P} sees $E>0$ along filaments (selected from intensity data), but no $B$-mode signal is above the noise.  We predict that it should be there in higher fidelity data.}
\revisit{A stacking analysis of Planck data \citep{2016A&A...586A.141P} sees $E>0$ along filaments (selected from intensity data), but claim no $B$-mode signal. We predict that a B-mode signal from filaments should be present, since they have a finite length and a varying radius of curvature. A naive filament stacking analysis is only suited for making E-mode detections. Detecting B-modes from filament edges will require a more careful edge stacked filament analysis. While the detectability of B-mode signature from filaments from Planck data calls for a more careful assessment, it should be detectable in higher fidelity data.}

\revisit{In \cite{Zaldarriaga2001a} it is argued that linear filaments of infinite length in which the configuration of the polarization vector is parallel or perpendicular to the length of the filament cannot produce B-mode power. While this conclusion is true for filaments of infinite length, in reality we expect filaments that are observed on the sky to have a finite length and in this case one expects a definite B-mode pattern from the edges of the filament. One also does not expect that filaments have a constant radius of curvature and this will result in an additional generation of B-mode power. Therefore our study indicates results which subtly differ from those presented in \cite{Zaldarriaga2001a} and these differences can be attributed to the more realistic assumptions about the geometry of the filaments made in our analysis.} \comment{Move to conclusions?}

>>>>

The $B$ part of the Green's function, traced along the filament, cancels itself except at the filament ends.  This results in a non-zero $B$ pattern for the filament.  For a North--South filament, the $B$-mode pattern is positive on the North-East and South-West, and negative in the North-West and South-East.  The size of this $B$-mode pattern is set by the dimensions of the filament, chiefly the width.

@@

Changing the polarization direction within the filament changes the $E/B$ patterns.  A $90^\circ$ rotation of the polarization with respect to the filament changes the sign of both $E$ and $B$.  In a polarization pattern aligned at $45^\circ$ to the filament, the $E$ pattern will swap with the $B$ pattern.  The work in \cite{Zaldarriaga2001a} correctly argued that linear filaments (infinite and without ends) can only produce $E$-power, but we find that for realistic filaments, the ends produce $B$-mode power with a clear signature.  Careful study of the $E/B$ mode power in filaments can provide insights into the orientation of the magnetic field with respect to the length of the filament.

The intuition from the real-space kernels holds also when we distort the shape of the filament.  If the filament were bent around into a circle, the positive and negative parts of the $B$ pattern will cancel, and we are left with a hoop of pure $E$ pattern. Here it is important to note that this cancellation will not happen for an ellipse or any loop of non-contant radius of curvature.  The same general description holds for a spiral-shaped filament, which can can be viewed as distortion of the straight filament.  The filament is highlighted by positive $E>0$.  The $E$-pattern is more negative on the interior of a curve than on the exterior, and the concentric rings of filamentary structure make an increasingly negative $E$ value inside. While the $B$-pattern is again concentrated at the ends of the filament in an oriented pair of positive/negative fluctuations, note that it is non-vanishing in the intermediate regions along the filament due to the varying radius of curvature of the spiral.
%\footnote{ \revisit{Why do we use this specific spiral - Its easy to compute the tangent to the spiral at any point.}}.

%\comment{If this is correct, then one expects to see a higher B-mode signal in intermediate regions of the spiral, as compared to the B-mode signal in the intermediate regions of the linear filament. This  then should be more visible in a logarithmically scaled version of Fig. 8.}  % K: addressed with new plot

%\revisit{Note that we have made the simplifying assumption that the polarization headless vectors are perpendicular to the magnetic field which is always assumed parallel to the tangent to the filament at any given point. This results in generating dominantly an E-mode signal while the B-mode signal is present only near the edges. By similar arguments it can be shown that that polarization pattern will be dominated by B-modes if the polarization vectors were aligned at $+45$ (or $-45$) degrees to the tangent to the filament while the E-mode signal would only be present at the edges. One way to get this configuration is to change the orientation of the magnetic field vector with respect to the tangent to the filament. Therefore a careful study of the $E/B$ mode power in filaments can provide insights into the orientation of the magnetic field with respect to the length of the filament.} % K: addressed

%\revisit{A stacking analysis of Planck data \citep{2016A&A...586A.141P} sees $E>0$ along filaments (selected from intensity data), but no $B$-mode signal is above the noise.  We predict that it should be there in higher fidelity data.}
A stacking analysis of Planck data \citep{2016A&A...586A.141P} sees $E>0$ along filaments (selected from intensity data), but claim no $B$-mode signal above the noise. We predict that a B-mode signal from filaments should be present, since they have a finite length of a few degrees.  Detecting B-modes from filaments will be easier with more signal-to-noise and may require a more careful filament analysis, rescaling and aligning the filament ends.  The detectability of the $B$-mode signature from stacked filaments in Planck data calls for a more careful assessment, we should see it in higher-fidelity data.

%\revisit{In \cite{Zaldarriaga2001a} it is argued that linear filaments of infinite length in which the configuration of the polarization vector is parallel or perpendicular to the length of the filament cannot produce B-mode power. While this conclusion is true for filaments of infinite length, in reality we expect filaments that are observed on the sky to have a finite length and in this case one expects a definite B-mode pattern from the edges of the filament. One also does not expect that filaments have a constant radius of curvature and this will result in an additional generation of B-mode power. Therefore our study indicates results which subtly differ from those presented in \cite{Zaldarriaga2001a} and these differences can be attributed to the more realistic assumptions about the geometry of the filaments made in our analysis.} \comment{Move to conclusions?}
